---
title: "Appendix: Exploratory analysis, survey non-response and imputation"
author: "Frank Edwards"
date: "`r Sys.Date()`"
output: 
  bookdown::pdf_document2:
    toc: false
    number_sections: false
---

IMPUTATION:

ID CORE PREDICTORS - set in one block

FOR EACH INSTRUMENT, ID VARIABLES NEEDED FOR INDEX
ID OUTCOMES FOR MODELS 

EXCLUDE OTHERS

OUTCOMES LISTED IN PREREG: 
https://clinicaltrials.gov/study/NCT05857553#study-plan

SF36 
K10
PSS4
CHAOS
PSS
FoodInsec
MissedCare
Truancy
RelativeChildCare
NonRelChildCare
CenterCare
CFPB FIN SCALE

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F)
library(mice)
library(miceadds)
library(micemd)
library(Amelia)
library(tidyverse)

dat<-read_csv("./data/cgir_NY_ocfs_FEClean04225.csv")
# define variable blocks
varBlocks=list(
  demogs=names(dat)[1:24],
  sf=names(dat)[25:43],
  k10=names(dat)[44:54],
  ps=names(dat)[55:58],
  fin=names(dat)[60:129],
  hsg=names(dat)[130:146],
  chaos=names(dat)[147:161],
  am=names(dat)[162:185],
  hhfi=names(dat)[186:190],
  ah=names(dat)[191:202],
  fw=names(dat)[203:212],
  ins=names(dat)[213:215],
  med=names(dat)[216:222],
  scl=names(dat)[223:228],
  chld=names(dat)[229:246],
  pss=names(dat)[247:264]
)

imp_init<-mice(dat,
           blocks = varBlocks,
           maxit = 0)
```





Trial 1: subset to sf36 + demographics

How do I want to do this: maybe easier to explicitly select only those columns to be imputed rather than start with everything. ID a set of core static predictors in xwalk, then work through each instrument / block. 


## DEMOGRAPHIC PREDICTORS

```{r}


## retain excluded as characters, convert others to factor / numeric
dat<-dat %>% 
  mutate(GROUP = factor(GROUP),
         County = factor(County),
         Wave = factor(Wave),
         Gender = factor(Gender),
         Ethnicity = factor(Ethnicity),
         MaritalStat = factor(MaritalStat),
         Race = factor(Race),
         Edu = factor(Edu),
         LangHome = factor(LangHome),
         SexOrient = factor(
           case_when(
             SexOrient == "Straight or Heterosexual" ~ "Straight or Heterosexual",
             SexOrient != "Straight or Heterosexual" ~ "Not Straight or Heterosexual"
           )))
## then subset to w1 for demogs, effectively a deterministic imputation
 
```

## trial 1 sf36

```{r}
temp_demog<-dat %>% 
  select(varBlocks$demogs) %>% 
  filter(Wave == 1) %>% 
  select(-Wave) 

# pivot data to wide
temp_sf<-dat %>% 
  select(PIN, Wave, varBlocks$sf) %>% 
  pivot_wider(id_cols = PIN, names_from = Wave, names_sep = "_W",
              values_from = SF_HealthGen:SF_HealthState_4)

### join to demog
temp_sf<-temp_demog %>% 
  left_join(temp_sf)

## initiate mids object
sf_imp<-mice(temp_sf, maxit = 0)
# disable vars to be excluded
sf_predmat<-sf_imp$predictorMatrix
# ID POSITION of vars
excluded<-c("PIN", "InWave", "SurveyStart", "DOB", 
            "PartnerStat", "PartnerGender", "EduOther",
            "LangOther")
excluded_pos<-which(names(temp_sf)%in%excluded)
# # switch them off
# sf_predmat[, excluded_pos]<-0
# # methods
# sf_meth<-sf_imp$method
# # default is pmm for numeric, polyreg for categorical, logreg for binary. that's good here
# sf_imp<-mice(temp_sf,
#              m = 3, 
#              method = sf_meth,
#              predictorMatrix = sf_predmat)

## problems: gender other; langhome tagalog; eduElemetnetary thru 5
# race: combine HPI + ASIAN
# small cell sizes there. collapse
temp_demog<-temp_demog %>% 
  mutate(
    Gender = factor(case_when(
      Gender == "Male" ~ "Male",
      Gender != "Male" ~ "Female or other")),
    LangHome = factor(case_when(
      LangHome == "English" ~ "English",
      LangHome != "English" ~ "Not English"
    )),
    Edu = factor(case_when(
      Edu == "Elementary school (through Grade 5)" | 
        Edu == "No formal education" ~ "Low or no formal education",
      T ~ Edu
    )),
    Race = factor(case_when(
      Race == "HPI" ~ "API",
      Race == "Asian" ~ "API",
      T ~ Race
    ))
  )

# try again

# pivot data to wide
temp_sf<-dat %>% 
  select(PIN, Wave, varBlocks$sf) %>% 
  pivot_wider(id_cols = PIN, names_from = Wave, names_sep = "_W",
              values_from = SF_HealthGen:SF_HealthState_4)

### join to demog
temp_sf<-temp_demog %>% 
  left_join(temp_sf)

## initiate mids object
sf_imp<-mice(temp_sf, maxit = 0)
# disable vars to be excluded
sf_predmat<-sf_imp$predictorMatrix
# ID POSITION of vars
excluded<-c("PIN", "InWave", "SurveyStart", "DOB", 
            "PartnerStat", "PartnerGender", "EduOther",
            "LangOther")
excluded_pos<-which(names(temp_sf)%in%excluded)
# switch them off
sf_predmat[, excluded_pos]<-0
# methods, use pmm for all, logreg and polyreg aren't working well
sf_meth<-sf_imp$method
sf_meth[which(!names(sf_meth)%in%excluded)]<-"pmm"
# default is pmm for numeric, polyreg for categorical, logreg for binary. that's good here
sf_imp<-mice(temp_sf,
             m = 5, 
             method = sf_meth,
             predictorMatrix = sf_predmat)

```

observed + posterior densitites of demographic variables

tables for binaries and some categoricals, densities for continuous

```{r}
# imputed vars
sf_imp$method[sf_imp$method=="pmm"]
temp<-mice::complete(sf_imp, action = "long", include = T)
```

```{r}
# gender
temp %>% 
  group_by(.imp, Gender) %>% 
  summarize(n = n()) %>% 
  mutate(prop = n / sum(n)) %>% 
  ggplot(aes(color = .imp==0, y = prop, group = .imp,
             x = Gender)) + 
  geom_point(position = position_dodge(width = 0.3))
```

minimal variation on gender

```{r}
densityplot(sf_imp,
            ~ HH_AdultsAll + HH_AdultsSome)
```

that looks fine

```{r}
temp %>% 
  group_by(.imp, MaritalStat) %>% 
  summarize(n = n()) %>% 
  mutate(prop = n / sum(n))%>% 
  ggplot(aes(color = .imp==0, y = prop, group = .imp,
             x = MaritalStat)) + 
  geom_point(position = position_dodge(width = 0.3))
```

also looks fine

```{r}
temp %>% 
  group_by(.imp, Ethnicity, Race) %>% 
  summarize(n = n()) %>% 
  mutate(prop = n / sum(n))%>% 
  ggplot(aes(color = .imp==0, x = n, group = .imp,
             y = Race)) + 
  geom_point(position = position_dodge(width = 0.3)) + 
  facet_wrap(~Ethnicity)
```

looks fine - education

```{r}
temp %>% 
  group_by(.imp, Edu) %>% 
  summarize(n = n()) %>% 
  mutate(prop = n / sum(n))%>% 
  ggplot(aes(color = .imp==0, x = prop, group = .imp,
             y = Edu)) + 
  geom_point(position = position_dodge(width = 0.4),
             alpha = 0.5)
```

langhome

```{r}
temp %>% 
  group_by(.imp, LangHome) %>% 
  summarize(n = n()) %>% 
  mutate(prop = n / sum(n))%>% 
  ggplot(aes(color = .imp==0, x = prop, group = .imp,
             y = LangHome)) + 
  geom_point(position = position_dodge(width = 0.4),
             alpha = 0.5)
```

sexorient

```{r}
temp %>% 
  group_by(.imp, SexOrient) %>% 
  summarize(n = n()) %>% 
  mutate(prop = n / sum(n))%>% 
  ggplot(aes(color = .imp==0, x = prop, group = .imp,
             y = SexOrient)) + 
  geom_point(position = position_dodge(width = 0.4),
             alpha = 0.5)
```

sf healthgen

```{r}
densityplot(sf_imp,
            ~ SF_HealthGen_W1 + SF_HealthGen_W2 + 
              SF_HealthGen_W3 + SF_HealthGen_W4)
```

sf healthlim1:10

```{r}
densityplot(sf_imp,
            ~ SF_HealthLim_1_W1 + SF_HealthLim_1_W2 + 
              SF_HealthLim_1_W3 + SF_HealthLim_1_W4 + 
              SF_HealthLim_2_W1 + SF_HealthLim_2_W2 + 
              SF_HealthLim_2_W3 + SF_HealthLim_2_W4 + 
              SF_HealthLim_3_W1 + SF_HealthLim_3_W2 + 
              SF_HealthLim_3_W3 + SF_HealthLim_3_W4 + 
              SF_HealthLim_4_W1 + SF_HealthLim_4_W2 + 
              SF_HealthLim_4_W3 + SF_HealthLim_4_W4,
            layout = c(4,4))
```

```{r}
densityplot(sf_imp,
            ~ SF_HealthLim_5_W1 + SF_HealthLim_5_W2 + 
              SF_HealthLim_5_W3 + SF_HealthLim_5_W4 + 
              SF_HealthLim_6_W1 + SF_HealthLim_6_W2 + 
              SF_HealthLim_6_W3 + SF_HealthLim_6_W4 + 
              SF_HealthLim_7_W1 + SF_HealthLim_7_W2 + 
              SF_HealthLim_7_W3 + SF_HealthLim_7_W4 + 
              SF_HealthLim_8_W1 + SF_HealthLim_8_W2 + 
              SF_HealthLim_8_W3 + SF_HealthLim_8_W4,
            layout = c(4,4))
```


```{r}
densityplot(sf_imp,
            ~ SF_HealthLim_9_W1 + SF_HealthLim_9_W2 + 
              SF_HealthLim_9_W3 + SF_HealthLim_9_W4 + 
              SF_HealthLim_10_W1 + SF_HealthLim_10_W2 + 
              SF_HealthLim_10_W3 + SF_HealthLim_10_W4,
            layout = c(2,4))
```


```{r}
densityplot(sf_imp,
            ~ SF_Phys_1_W1 + SF_Phys_1_W2 + 
              SF_Phys_1_W3 + SF_Phys_1_W4 + 
              SF_Phys_2_W1 + SF_Phys_2_W2 + 
              SF_Phys_2_W3 + SF_Phys_2_W4 + 
              SF_Phys_3_W1 + SF_Phys_3_W2 + 
              SF_Phys_3_W3 + SF_Phys_3_W4 + 
              SF_Phys_4_W1 + SF_Phys_4_W2 + 
              SF_Phys_4_W3 + SF_Phys_4_W4,
            layout = c(4, 4))
```

health state

```{r}
densityplot(sf_imp,
            ~ SF_HealthState_1_W1 + SF_HealthState_1_W2 + 
              SF_HealthState_1_W3 + SF_HealthState_1_W4 + 
              SF_HealthState_2_W1 + SF_HealthState_2_W2 + 
              SF_HealthState_2_W3 + SF_HealthState_2_W4 + 
              SF_HealthState_3_W1 + SF_HealthState_3_W2 + 
              SF_HealthState_3_W3 + SF_HealthState_3_W4 + 
              SF_HealthState_4_W1 + SF_HealthState_4_W2 + 
              SF_HealthState_4_W3 + SF_HealthState_4_W4,
            layout = c(4,4))
```

## Kessler 10

```{r}
# pivot data to wide
temp_k10<-temp_demog %>% 
  left_join(
    dat %>% 
      select(PIN, Wave, varBlocks$k10) %>% 
      pivot_wider(id_cols = PIN, names_from = Wave, names_sep = "_W",
                  values_from = K10_1:K10phys))
```

Impute

```{r}
## initiate mids object
k10_imp<-mice(temp_k10, maxit = 0)
# disable vars to be excluded
k10_predmat<-k10_imp$predictorMatrix
# ID POSITION of vars
excluded<-c("PIN", "InWave", "SurveyStart", "DOB", 
            "PartnerStat", "PartnerGender", "EduOther",
            "LangOther")
excluded_pos<-which(names(temp_sf)%in%excluded)
# switch them off
k10_predmat[, excluded_pos]<-0
# methods, use pmm for all, logreg and polyreg aren't working well
k10_meth<-k10_imp$method
k10_meth[which(!names(k10_meth)%in%excluded)]<-"pmm"
# default is pmm for numeric, polyreg for categorical, logreg for binary. that's good here
k10_imp<-mice(temp_k10,
             m = 5, 
             method = k10_meth,
             predictorMatrix = k10_predmat)
```

Evaluate diagnostics

```{r}
densityplot(k10_imp,
            ~ K10_1_W1 + K10_1_W2 + 
              K10_1_W3 + K10_1_W4 + 
              K10_2_W1 + K10_2_W2 + 
              K10_2_W3 + K10_2_W4 + 
              K10_3_W1 + K10_3_W2 + 
              K10_3_W3 + K10_3_W4 + 
              K10_4_W1 + K10_4_W2 + 
              K10_4_W3 + K10_4_W4,
            layout = c(4, 4))
```


```{r}
densityplot(k10_imp,
            ~ K10_5_W1 + K10_5_W2 + 
              K10_5_W3 + K10_5_W4 + 
              K10_6_W1 + K10_6_W2 + 
              K10_6_W3 + K10_6_W4 + 
              K10_7_W1 + K10_7_W2 + 
              K10_7_W3 + K10_7_W4 + 
              K10_8_W1 + K10_8_W2 + 
              K10_8_W3 + K10_8_W4,
            layout = c(4, 4))
```

```{r}
densityplot(k10_imp,
            ~ K10_9_W1 + K10_9_W2 + 
              K10_9_W3 + K10_9_W4 + 
              K10_10_W1 + K10_10_W2 + 
              K10_10_W3 + K10_10_W4 + 
              K10phys_W1 + K10phys_W2 + 
              K10phys_W3 + K10phys_W4,
            layout = c(4, 3))
```

## ps

```{r}
# pivot data to wide
temp_ps<-temp_demog %>% 
  left_join(
    dat %>% 
      select(PIN, Wave, varBlocks$ps) %>% 
      pivot_wider(id_cols = PIN, names_from = Wave, names_sep = "_W",
                  values_from = PS_2:PS_10))

## initiate mids object
ps_imp<-mice(temp_ps, maxit = 0)
# disable vars to be excluded
ps_predmat<-ps_imp$predictorMatrix
# ID POSITION of vars
excluded<-c("PIN", "InWave", "SurveyStart", "DOB", 
            "PartnerStat", "PartnerGender", "EduOther",
            "LangOther")
excluded_pos<-which(names(temp_sf)%in%excluded)
# switch them off
ps_predmat[, excluded_pos]<-0
# methods, use pmm for all, logreg and polyreg aren't working well
ps_meth<-ps_imp$method
ps_meth[which(!names(ps_meth)%in%excluded)]<-"pmm"
# default is pmm for numeric, polyreg for categorical, logreg for binary. that's good here
ps_imp<-mice(temp_ps,
             m = 5, 
             method = ps_meth,
             predictorMatrix = ps_predmat)
```

visualize

```{r}
densityplot(ps_imp,
            ~ PS_2_W1 + PS_2_W2 + PS_2_W3 + PS_2_W4 + 
              PS_4_W1 + PS_4_W2 + PS_4_W3 + PS_4_W4 + 
              PS_5_W1 + PS_5_W2 + PS_5_W3 + PS_5_W4 + 
              PS_10_W1 + PS_10_W2 + PS_10_W3 + PS_10_W4,
            layout = c(4, 4))
```

looks fine - variance in w1 looks like small prop missing to me

## Housing

```{r}
# pivot data to wide
# do needed cleaning for small cells etc
temp_hsg<-temp_demog %>% 
  left_join(
    dat %>% 
      select(PIN, Wave, varBlocks$hsg) %>% 
      mutate(Hsg_Stat = ifelse(Hsg_Stat == "Rent-to-own", "Renter", Hsg_Stat)) %>% 
      mutate(across(c(Hsg_Stat, Hsg_Qual, Hsg_NeighQual), as.factor)) %>% 
      pivot_wider(id_cols = PIN, names_from = Wave, names_sep = "_W",
                  values_from = Hsg_Stat:Hsg_MortDefault))

## initiate mids object
hsg_imp<-mice(temp_hsg, maxit = 0)
# disable vars to be excluded
hsg_predmat<-hsg_imp$predictorMatrix
# ID POSITION of vars
excluded<-c("PIN", "InWave", "SurveyStart", "DOB", 
            "PartnerStat", "PartnerGender", "EduOther",
            "LangOther", "Hsg_StatOther",
            "RentAssist", "Hsg_Current", "Util")
# ID position of excludeds
excludeds<-paste(excluded, collapse = "|")
excluded_pos<-grep(excludeds, names(temp_hsg))

# switch them off
hsg_predmat[, excluded_pos]<-0
# methods, use pmm for all, logreg and polyreg aren't working well
hsg_meth<-hsg_imp$method
hsg_meth[which(!names(hsg_meth)%in%excluded)]<-"pmm"
hsg_meth[excluded_pos]<-""
# default is pmm for numeric, polyreg for categorical, logreg for binary. that's good here
hsg_imp<-mice(temp_hsg,
             m = 5, 
             method = hsg_meth,
             predictorMatrix = hsg_predmat)
```

visualize

```{r}
densityplot(hsg_imp,
            ~ Hsg_Stat_W1 + Hsg_Stat_W2 + Hsg_Stat_W3 + Hsg_Stat_W4 + 
              Hsg_Move_W1 + Hsg_Move_W2 + Hsg_Move_W3 + Hsg_Move_W4 + 
              Hsg_Qual_W1 + Hsg_Qual_W2 + Hsg_Qual_W3 + Hsg_Qual_W4 + 
              Hsg_NeighQual_W1 + Hsg_NeighQual_W2 + Hsg_NeighQual_W3 + Hsg_NeighQual_W4,
            layout = c(4, 4))
```

```{r}
densityplot(hsg_imp,
            ~ Hsg_RentAmt_W1 + Hsg_RentAmt_W2 + Hsg_RentAmt_W3 + Hsg_RentAmt_W4 + 
              Hsg_Evict_W1 + Hsg_Evict_W2 + Hsg_Evict_W3 + Hsg_Evict_W4 + 
              Hsg_MortDefault_W1 + Hsg_MortDefault_W2 + Hsg_MortDefault_W3 + Hsg_MortDefault_W4,
            layout = c(4, 3))
```

## CHAOS

```{r}
# pivot data to wide
temp_Chaos<-temp_demog %>% 
  left_join(
    dat %>% 
      select(PIN, Wave, varBlocks$chaos) %>% 
      pivot_wider(id_cols = PIN, names_from = Wave, names_sep = "_W",
                  values_from = Chaos_1:Chaos_15))

## initiate mids object
Chaos_imp<-mice(temp_Chaos, maxit = 0)
# disable vars to be excluded
Chaos_predmat<-Chaos_imp$predictorMatrix
# ID POSITION of vars
excluded<-c("PIN", "InWave", "SurveyStart", "DOB", 
            "PartnerStat", "PartnerGender", "EduOther",
            "LangOther")
excluded_pos<-which(names(temp_sf)%in%excluded)
# switch them off
Chaos_predmat[, excluded_pos]<-0
# methods, use pmm for all, logreg and polyreg aren't working well
Chaos_meth<-Chaos_imp$method
Chaos_meth[which(!names(Chaos_meth)%in%excluded)]<-"pmm"
# default is pmm for numeric, polyreg for categorical, logreg for binary. that's good here
Chaos_imp<-mice(temp_Chaos,
             m = 5, 
             method = Chaos_meth,
             predictorMatrix = Chaos_predmat)
```

visualize

```{r}
densityplot(Chaos_imp,
            ~ Chaos_1_W1 + Chaos_1_W2 + Chaos_1_W3 + Chaos_1_W4 + 
              Chaos_2_W1 + Chaos_2_W2 + Chaos_2_W3 + Chaos_2_W4 + 
              Chaos_3_W1 + Chaos_3_W2 + Chaos_3_W3 + Chaos_3_W4 + 
              Chaos_4_W1 + Chaos_4_W2 + Chaos_4_W3 + Chaos_4_W4,
            layout = c(4, 4))
```

```{r}
densityplot(Chaos_imp,
            ~ Chaos_5_W1 + Chaos_5_W2 + Chaos_5_W3 + Chaos_5_W4 + 
              Chaos_6_W1 + Chaos_6_W2 + Chaos_6_W3 + Chaos_6_W4 + 
              Chaos_7_W1 + Chaos_7_W2 + Chaos_7_W3 + Chaos_7_W4 + 
              Chaos_8_W1 + Chaos_8_W2 + Chaos_8_W3 + Chaos_8_W4,
            layout = c(4, 4))
```

```{r}
densityplot(Chaos_imp,
            ~ Chaos_9_W1 + Chaos_9_W2 + Chaos_9_W3 + Chaos_9_W4 + 
              Chaos_10_W1 + Chaos_10_W2 + Chaos_10_W3 + Chaos_10_W4 + 
              Chaos_11_W1 + Chaos_11_W2 + Chaos_11_W3 + Chaos_11_W4 + 
              Chaos_12_W1 + Chaos_12_W2 + Chaos_12_W3 + Chaos_12_W4,
            layout = c(4, 4))
```

```{r}
densityplot(Chaos_imp,
            ~ Chaos_13_W1 + Chaos_13_W2 + Chaos_13_W3 + Chaos_13_W4 + 
              Chaos_14_W1 + Chaos_14_W2 + Chaos_14_W3 + Chaos_14_W4 + 
              Chaos_15_W1 + Chaos_15_W2 + Chaos_15_W3 + Chaos_15_W4, 
            layout = c(4, 3))
```

## Adult mattering

```{r}
# pivot data to wide
temp_am<-temp_demog %>% 
  left_join(
    dat %>% 
      select(PIN, Wave, varBlocks$am) %>% 
      pivot_wider(id_cols = PIN, names_from = Wave, names_sep = "_W",
                  values_from = AM_1:AM_24))

## initiate mids object
am_imp<-mice(temp_am, maxit = 0)
# disable vars to be excluded
am_predmat<-am_imp$predictorMatrix
# ID POSITION of vars
excluded<-c("PIN", "InWave", "SurveyStart", "DOB", 
            "PartnerStat", "PartnerGender", "EduOther",
            "LangOther")
excluded_pos<-which(names(temp_sf)%in%excluded)
# switch them off
am_predmat[, excluded_pos]<-0
# methods, use pmm for all, logreg and polyreg aren't working well
am_meth<-am_imp$method
am_meth[which(!names(am_meth)%in%excluded)]<-"pmm"
# default is pmm for numeric, polyreg for categorical, logreg for binary. that's good here
am_imp<-mice(temp_am,
             m = 5, 
             method = am_meth,
             predictorMatrix = am_predmat)
```

visualize

```{r}
densityplot(am_imp,
            ~ AM_1_W1 + AM_1_W2 + AM_1_W3 + AM_1_W4 + 
              AM_2_W1 + AM_2_W2 + AM_2_W3 + AM_2_W4 + 
              AM_3_W1 + AM_3_W2 + AM_3_W3 + AM_3_W4 + 
              AM_4_W1 + AM_4_W2 + AM_4_W3 + AM_4_W4,
            layout = c(4, 4))
```

```{r}
densityplot(am_imp,
            ~ AM_5_W1 + AM_5_W2 + AM_5_W3 + AM_5_W4 + 
              AM_6_W1 + AM_6_W2 + AM_6_W3 + AM_6_W4 + 
              AM_7_W1 + AM_7_W2 + AM_7_W3 + AM_7_W4 + 
              AM_8_W1 + AM_8_W2 + AM_8_W3 + AM_8_W4,
            layout = c(4, 4))
```

```{r}
densityplot(am_imp,
            ~ AM_9_W1 + AM_9_W2 + AM_9_W3 + AM_9_W4 + 
              AM_10_W1 + AM_10_W2 + AM_10_W3 + AM_10_W4 + 
              AM_11_W1 + AM_11_W2 + AM_11_W3 + AM_11_W4 + 
              AM_12_W1 + AM_12_W2 + AM_12_W3 + AM_12_W4,
            layout = c(4, 4))
```

```{r}
densityplot(am_imp,
            ~ AM_13_W1 + AM_13_W2 + AM_13_W3 + AM_13_W4 + 
              AM_14_W1 + AM_14_W2 + AM_14_W3 + AM_14_W4 + 
              AM_15_W1 + AM_15_W2 + AM_15_W3 + AM_15_W4 + 
              AM_16_W1 + AM_16_W2 + AM_16_W3 + AM_16_W4,
            layout = c(4, 4))
```

```{r}
densityplot(am_imp,
            ~ AM_17_W1 + AM_17_W2 + AM_17_W3 + AM_17_W4 + 
              AM_18_W1 + AM_18_W2 + AM_18_W3 + AM_18_W4 + 
              AM_19_W1 + AM_19_W2 + AM_19_W3 + AM_19_W4 + 
              AM_20_W1 + AM_20_W2 + AM_20_W3 + AM_20_W4,
            layout = c(4, 4))
```

```{r}
densityplot(am_imp,
            ~ AM_21_W1 + AM_21_W2 + AM_21_W3 + AM_21_W4 + 
              AM_22_W1 + AM_22_W2 + AM_22_W3 + AM_22_W4 + 
              AM_23_W1 + AM_23_W2 + AM_23_W3 + AM_23_W4 + 
              AM_24_W1 + AM_24_W2 + AM_24_W3 + AM_24_W4,
            layout = c(4, 4))
```

## PSS

```{r}
# pivot data to wide
temp_ps<-temp_demog %>% 
  left_join(
    dat %>% 
      select(PIN, Wave, varBlocks$ps) %>% 
      pivot_wider(id_cols = PIN, names_from = Wave, names_sep = "_W",
                  values_from = PS_2:PS_10))

## initiate mids object
ps_imp<-mice(temp_ps, maxit = 0)
# disable vars to be excluded
ps_predmat<-ps_imp$predictorMatrix
# ID POSITION of vars
excluded<-c("PIN", "InWave", "SurveyStart", "DOB", 
            "PartnerStat", "PartnerGender", "EduOther",
            "LangOther")
excluded_pos<-which(names(temp_sf)%in%excluded)
# switch them off
ps_predmat[, excluded_pos]<-0
# methods, use pmm for all, logreg and polyreg aren't working well
ps_meth<-ps_imp$method
ps_meth[which(!names(ps_meth)%in%excluded)]<-"pmm"
# default is pmm for numeric, polyreg for categorical, logreg for binary. that's good here
ps_imp<-mice(temp_ps,
             m = 5, 
             method = ps_meth,
             predictorMatrix = ps_predmat)
```

visualize

```{r}
densityplot(ps_imp,
            ~ PS_2_W1 + PS_2_W2 + PS_2_W3 + PS_2_W4 + 
              PS_4_W1 + PS_4_W2 + PS_4_W3 + PS_4_W4 + 
              PS_5_W1 + PS_5_W2 + PS_5_W3 + PS_5_W4 + 
              PS_10_W1 + PS_10_W2 + PS_10_W3 + PS_10_W4,
            layout = c(4, 4))
```

## FoodInsec

```{r}
# pivot data to wide
temp_hhfi<-temp_demog %>% 
  left_join(
    dat %>% 
      select(PIN, Wave, varBlocks$hhfi) %>% 
      pivot_wider(id_cols = PIN, names_from = Wave, names_sep = "_W",
                  values_from = HHFI_Insuf:HHFI_Util))

## initiate mids object
hhfi_imp<-mice(temp_hhfi, maxit = 0)
# disable vars to be excluded
hhfi_predmat<-hhfi_imp$predictorMatrix
# ID POSITION of vars
excluded<-c("PIN", "InWave", "SurveyStart", "DOB", 
            "PartnerStat", "PartnerGender", "EduOther",
            "LangOther")
excluded_pos<-which(names(temp_sf)%in%excluded)
# switch them off
hhfi_predmat[, excluded_pos]<-0
# methods, use pmm for all, logreg and polyreg aren't working well
hhfi_meth<-hhfi_imp$method
hhfi_meth[which(!names(hhfi_meth)%in%excluded)]<-"pmm"
# default is pmm for numeric, polyreg for categorical, logreg for binary. that's good here
hhfi_imp<-mice(temp_hhfi,
             m = 5, 
             method = hhfi_meth,
             predictorMatrix = hhfi_predmat)
```

visualize

```{r}
## use dot plot for binary proportions
temp<-mice::complete(hhfi_imp, action = "long", include = T)

temp %>% 
  group_by(.imp) %>% 
  summarize(across(HHFI_Insuf_W1:HHFI_Util_W4, \(x) sum(x, na.rm=T)/ sum(!(is.na(x))))) %>% 
  pivot_longer(cols = HHFI_Insuf_W1:HHFI_Util_W4, 
               names_to = "var", 
               values_to = "prop") %>% 
  ggplot(aes(color = .imp==0, y = var,
             x = prop)) + 
  geom_boxplot()
```

## Adult hope

```{r}
# pivot data to wide
temp_ah<-temp_demog %>% 
  left_join(
    dat %>% 
      select(PIN, Wave, varBlocks$ah) %>% 
      pivot_wider(id_cols = PIN, names_from = Wave, names_sep = "_W",
                  values_from = AH_1:AH_12))

## initiate mids object
ah_imp<-mice(temp_ah, maxit = 0)
# disable vars to be excluded
ah_predmat<-ah_imp$predictorMatrix
# ID POSITION of vars
excluded<-c("PIN", "InWave", "SurveyStart", "DOB", 
            "PartnerStat", "PartnerGender", "EduOther",
            "LangOther")
excluded_pos<-which(names(temp_sf)%in%excluded)
# switch them off
ah_predmat[, excluded_pos]<-0
# methods, use pmm for all, logreg and polyreg aren't working well
ah_meth<-ah_imp$method
ah_meth[which(!names(ah_meth)%in%excluded)]<-"pmm"
# default is pmm for numeric, polyreg for categorical, logreg for binary. that's good here
ah_imp<-mice(temp_ah,
             m = 5, 
             method = ah_meth,
             predictorMatrix = ah_predmat)
```

visualize

```{r}
densityplot(ah_imp,
            ~ AH_1_W1 + AH_1_W2 + AH_1_W3 + AH_1_W4 + 
              AH_2_W1 + AH_2_W2 + AH_2_W3 + AH_2_W4 + 
              AH_3_W1 + AH_3_W2 + AH_3_W3 + AH_3_W4 + 
              AH_4_W1 + AH_4_W2 + AH_4_W3 + AH_4_W4,
            layout = c(4, 4))
```

```{r}
densityplot(ah_imp,
            ~ AH_5_W1 + AH_5_W2 + AH_5_W3 + AH_5_W4 + 
              AH_6_W1 + AH_6_W2 + AH_6_W3 + AH_6_W4 + 
              AH_7_W1 + AH_7_W2 + AH_7_W3 + AH_7_W4 + 
              AH_8_W1 + AH_8_W2 + AH_8_W3 + AH_8_W4,
            layout = c(4, 4))
```

```{r}
densityplot(ah_imp,
            ~ AH_9_W1 + AH_9_W2 + AH_9_W3 + AH_9_W4 + 
              AH_10_W1 + AH_10_W2 + AH_10_W3 + AH_10_W4 + 
              AH_11_W1 + AH_11_W2 + AH_11_W3 + AH_11_W4 + 
              AH_12_W1 + AH_12_W2 + AH_12_W3 + AH_12_W4,
            layout = c(4, 4))
```

## FW

```{r}
# pivot data to wide
temp_fw<-temp_demog %>% 
  left_join(
    dat %>% 
      select(PIN, Wave, varBlocks$fw) %>% 
      pivot_wider(id_cols = PIN, names_from = Wave, names_sep = "_W",
                  values_from = FW_1_1:FW_2_4))

## initiate mids object
fw_imp<-mice(temp_fw, maxit = 0)
# disable vars to be excluded
fw_predmat<-fw_imp$predictorMatrix
# ID POSITION of vars
excluded<-c("PIN", "InWave", "SurveyStart", "DOB", 
            "PartnerStat", "PartnerGender", "EduOther",
            "LangOther")
excluded_pos<-which(names(temp_sf)%in%excluded)
# switch them off
fw_predmat[, excluded_pos]<-0
# methods, use pmm for all, logreg and polyreg aren't working well
fw_meth<-fw_imp$method
fw_meth[which(!names(fw_meth)%in%excluded)]<-"pmm"
# default is pmm for numeric, polyreg for categorical, logreg for binary. that's good here
fw_imp<-mice(temp_fw,
             m = 5, 
             method = fw_meth,
             predictorMatrix = fw_predmat)
```

visualize

```{r}
densityplot(fw_imp,
            ~ FW_1_1_W1 + FW_1_1_W2 + FW_1_1_W3 + FW_1_1_W4 + 
              FW_1_2_W1 + FW_1_2_W2 + FW_1_2_W3 + FW_1_2_W4 + 
              FW_1_3_W1 + FW_1_3_W2 + FW_1_3_W3 + FW_1_3_W4 + 
              FW_1_4_W1 + FW_1_4_W2 + FW_1_4_W3 + FW_1_4_W4,
            layout = c(4, 4))
```

```{r}
densityplot(fw_imp,
            ~ FW_1_5_W1 + FW_1_5_W2 + FW_1_5_W3 + FW_1_5_W4 + 
              FW_1_6_W1 + FW_1_6_W2 + FW_1_6_W3 + FW_1_6_W4 + 
              FW_2_1_W1 + FW_2_1_W2 + FW_2_1_W3 + FW_2_1_W4 + 
              FW_2_2_W1 + FW_2_2_W2 + FW_2_2_W3 + FW_2_2_W4,
            layout = c(4, 4))
```

```{r}
densityplot(fw_imp,
            ~ FW_2_3_W1 + FW_2_3_W2 + FW_2_3_W3 + FW_2_3_W4 + 
              FW_2_4_W1 + FW_2_4_W2 + FW_2_4_W3 + FW_2_4_W4,
            layout = c(4, 2))
```

## Insurance and Medical

```{r}
# pivot data to wide
temp_ins<-temp_demog %>% 
  left_join(
    dat %>% 
      select(PIN, Wave, c(varBlocks$ins, varBlocks$med)) %>% 
      pivot_wider(id_cols = PIN, names_from = Wave, names_sep = "_W",
                  values_from = c(ParentInsurance:ChildInsuranceNo, MedicalAccess)))

## initiate mids object
ins_imp<-mice(temp_ins, maxit = 0)
# disable vars to be excluded
ins_predmat<-ins_imp$predictorMatrix
# ID POSITION of vars
excluded<-c("PIN", "InWave", "SurveyStart", "DOB", 
            "PartnerStat", "PartnerGender", "EduOther",
            "LangOther")
excluded_pos<-which(names(temp_sf)%in%excluded)
# switch them off
ins_predmat[, excluded_pos]<-0
# methods, use pmm for all, logreg and polyreg aren't working well
ins_meth<-ins_imp$method
ins_meth[which(!names(ins_meth)%in%excluded)]<-"pmm"
# default is pmm for numeric, polyreg for categorical, logreg for binary. that's good here
ins_imp<-mice(temp_ins,
             m = 5, 
             method = ins_meth,
             predictorMatrix = ins_predmat)
```

visualize

```{r}
## use dot plot for binary proportions
temp<-mice::complete(ins_imp, action = "long", include = T)

temp %>% 
  group_by(.imp) %>% 
  summarize(across(ParentInsurance_W1:MedicalAccess_W4, \(x) sum(x, na.rm=T)/ sum(!(is.na(x))))) %>% 
  pivot_longer(cols = ParentInsurance_W1:MedicalAccess_W4, 
               names_to = "var", 
               values_to = "prop") %>% 
  ggplot(aes(color = .imp==0, y = var,
             x = prop)) + 
  geom_boxplot()
```

## Truancy 

way too much missing data on child care to work with - going to ditch it. for w1, we've got 87% missing for center care

```{r}
### compute rowmeans, then impute the missings unless nchild = 0
dat<-dat %>% 
  rowwise() %>% 
  mutate(ChildMissSchool_mn = mean(c(
    ChildMissSchool1,ChildMissSchool2,
    ChildMissSchool2, ChildMissSchool3,
    ChildMissSchool4, ChildMissSchool5,
    ChildMissSchool6),
    na.rm=T))

# pivot data to wide
temp_scl<-temp_demog %>% 
  left_join(
    dat %>% 
      select(PIN, Wave, ChildMissSchool_mn, 
             childcare_relative_total, childcare_nonrelative_total,
             childcare_center_total) %>% 
      pivot_wider(id_cols = PIN, names_from = Wave, names_sep = "_W",
                  values_from = c(ChildMissSchool_mn:childcare_center_total)))

## initiate mids object
scl_imp<-mice(temp_scl, maxit = 0)
# disable vars to be excluded
scl_predmat<-scl_imp$predictorMatrix
# ID POSITION of vars
excluded<-c("PIN", "InWave", "SurveyStart", "DOB", 
            "PartnerStat", "PartnerGender", "EduOther",
            "LangOther")
excluded_pos<-which(names(temp_sf)%in%excluded)
# switch them off
scl_predmat[, excluded_pos]<-0
# methods, use pmm for all, logreg and polyreg aren't working well
scl_meth<-scl_imp$method
scl_meth[which(!names(scl_meth)%in%excluded)]<-"pmm"
# default is pmm for numeric, polyreg for categorical, logreg for binary. that's good here
scl_imp<-mice(temp_scl,
             m = 5, 
             method = scl_meth,
             predictorMatrix = scl_predmat)

```

missed school

```{r}
densityplot(scl_imp,
            ~ ChildMissSchool_mn_W1 + ChildMissSchool_mn_W2 + ChildMissSchool_mn_W3 + ChildMissSchool_mn_W4,
            layout = c(4,1))
```


## PSS

```{r}
# pivot data to wide
temp_pss<-temp_demog %>% 
  left_join(
    dat %>% 
      select(PIN, Wave, varBlocks$pss) %>% 
      pivot_wider(id_cols = PIN, names_from = Wave, names_sep = "_W",
                  values_from = ParentalStressScale_1:ParentalStressScale_18))

## initiate mids object
pss_imp<-mice(temp_pss, maxit = 0)
# disable vars to be excluded
pss_predmat<-pss_imp$predictorMatrix
# ID POSITION of vars
excluded<-c("PIN", "InWave", "SurveyStart", "DOB", 
            "PartnerStat", "PartnerGender", "EduOther",
            "LangOther")
excluded_pos<-which(names(temp_pss)%in%excluded)
# switch them off
pss_predmat[, excluded_pos]<-0
# methods, use pmm for all, logreg and polyreg aren't working well
pss_meth<-pss_imp$method
pss_meth[which(!names(pss_meth)%in%excluded)]<-"pmm"
# default is pmm for numeric, polyreg for categorical, logreg for binary. that's good here
pss_imp<-mice(temp_pss,
             m = 5, 
             method = ps_meth,
             predictorMatrix = ps_predmat)
```

visualize

```{r}
densityplot(pss_imp,
            ~ ParentalStressScale_1_W1 + ParentalStressScale_1_W2 + ParentalStressScale_1_W3 + ParentalStressScale_1_W4 + 
              ParentalStressScale_2_W1 + ParentalStressScale_2_W2 + ParentalStressScale_2_W3 + ParentalStressScale_2_W4 + 
              ParentalStressScale_3_W1 + ParentalStressScale_3_W2 + ParentalStressScale_3_W3 + ParentalStressScale_3_W4 + 
              ParentalStressScale_4_W1 + ParentalStressScale_4_W2 + ParentalStressScale_4_W3 + ParentalStressScale_4_W4,
            layout = c(4, 4))
```

```{r}
densityplot(pss_imp,
            ~ ParentalStressScale_5_W1 + ParentalStressScale_5_W2 + ParentalStressScale_5_W3 + ParentalStressScale_5_W4 + 
              ParentalStressScale_6_W1 + ParentalStressScale_6_W2 + ParentalStressScale_6_W3 + ParentalStressScale_6_W4 + 
              ParentalStressScale_7_W1 + ParentalStressScale_7_W2 + ParentalStressScale_7_W3 + ParentalStressScale_7_W4 + 
              ParentalStressScale_8_W1 + ParentalStressScale_8_W2 + ParentalStressScale_8_W3 + ParentalStressScale_8_W4,
            layout = c(4, 4))
```

```{r}
densityplot(pss_imp,
            ~ ParentalStressScale_9_W1 + ParentalStressScale_9_W2 + ParentalStressScale_9_W3 + ParentalStressScale_9_W4 + 
              ParentalStressScale_10_W1 + ParentalStressScale_10_W2 + ParentalStressScale_10_W3 + ParentalStressScale_10_W4 + 
              ParentalStressScale_11_W1 + ParentalStressScale_11_W2 + ParentalStressScale_11_W3 + ParentalStressScale_11_W4 + 
              ParentalStressScale_12_W1 + ParentalStressScale_12_W2 + ParentalStressScale_12_W3 + ParentalStressScale_12_W4,
            layout = c(4, 4))
```

```{r}
densityplot(pss_imp,
            ~ ParentalStressScale_13_W1 + ParentalStressScale_13_W2 + ParentalStressScale_13_W3 + ParentalStressScale_13_W4 + 
              ParentalStressScale_14_W1 + ParentalStressScale_14_W2 + ParentalStressScale_14_W3 + ParentalStressScale_14_W4 + 
              ParentalStressScale_15_W1 + ParentalStressScale_15_W2 + ParentalStressScale_15_W3 + ParentalStressScale_15_W4 + 
              ParentalStressScale_16_W1 + ParentalStressScale_16_W2 + ParentalStressScale_16_W3 + ParentalStressScale_16_W4,
            layout = c(4, 4))
```

```{r}
densityplot(pss_imp,
            ~ ParentalStressScale_17_W1 + ParentalStressScale_17_W2 + ParentalStressScale_17_W3 + ParentalStressScale_17_W4 + 
              ParentalStressScale_18_W1 + ParentalStressScale_18_W2 + ParentalStressScale_18_W3 + ParentalStressScale_18_W4,
            layout = c(4, 2))
```


## Joint imputation



## check relationships between high missingness and demogs / wave / group
